Effective collector efficiency of sparse and dense arrays of emergent cylindrical collectors in laminar-turbulent transitional flows

### 2020-07-22 drafts/outline.tex
working on 2.2

Particle capture in transitionally turbulent flows ($1<\Rey<1000$), which occur commonly in the environments and on the length scales of aquatic macrovegetation, does not follow the analytical expressions involving Reynolds number, collector diameter, and particle diameter ($d_p$), derived for creeping flow. Hence, empirical estimates must be used. \cite{Palmer_2004} introduced a power law expression for estimating capture efficiency of cylindrical collectors, which takes the form \[\eta=C{\Rey_c}^{a}R^{b}\,,\] where collector Reynolds number $\Rey_c=\frac{ud_c}{\nu}$, $u$ is flow velocity, $\nu$ is kinematic velocity, and $R=\frac{d_c}{d_p}$ is the ratio between collector diameter and particle diameter ($d_p$). $C$, $a$, and $b$ are empirically determined regression coefficients, the sign and magnitude of which may both vary depending on number of stems, stem surface properties, and other experimental parameters \citep[e.g.,][]{Palmer_2004, Fauria_2015}. It is worth noting that \cite{Palmer_2004} used $\eta$ in their notation for this model as opposed to $\eta\prime$ because their experiment involved an isolated single cylinder, coated with grease for which retention was assumed to be complete ($p_r = 1$).

As is the case with capture efficiency, turbulence is also theorized and empirically confirmed to affect the settling velocity of heavy, fine particles \citep{Nielsen_1993, Wang_2018}. Increasing turbulence is is one mechanism by which collectors could affect settling rates.

@@

For a fluid with a Rouse.

We used a treatment-control model to estimate effective collector efficiency at various densities of emergent collectors and various Reynold's numbers. Our model was based on that used in \cite{Fauria_2015}, where a more complete derivation can be found. To summarize, collection rate of a single cylinder is modeled, assuming steady, unidirectional flow, as $\bar{\phi}ud_ch_c$, where $\bar{\phi}$ is depth-averaged suspended particle concentration, and $h_c$ is collector height. For emergent collectors, all collectors have effective heights equal to water depth. \[\frac{d\bar{\phi}}{dt} = -[\frac{Cv_s}{h}(1-E_r) + \eta^{\prime}ud_cI_c]\bar{\phi} = -k\bar{\phi}\], % \[ is necessary instead of $$ because it preserves line numbers
and its which

>>>>

Particle capture in transitionally turbulent flows ($1<\Rey<1000$), which occur commonly in the environments and on the length scales of aquatic macrovegetation, does not follow the analytical expressions involving Reynolds number, collector diameter, and particle diameter ($d_p$), derived for creeping flow. Hence, empirical estimates must be used. \cite{Palmer_2004} introduced a power law expression for estimating capture efficiency of cylindrical collectors, which takes the form \[\eta=C{\Rey_c}^{a}R^{b}\,,\] where collector Reynolds number $\Rey_c=\frac{ud_c}{\nu}$, $u$ is flow velocity, $\nu$ is kinematic velocity, and $R=\frac{d_c}{d_p}$ is the ratio between collector diameter and particle diameter ($d_p$). $C$, $a$, and $b$ are empirically determined regression coefficients. It is worth noting that \cite{Palmer_2004} used $\eta$ in their notation for this model as opposed to $\eta\prime$ because their experiment involved an isolated single cylinder, coated with grease for which retention was assumed to be complete ($p_r = 1$).

As is the case with capture efficiency, turbulence is also theorized and empirically confirmed to affect the settling velocity of heavy, fine particles \citep{Nielsen_1993, Jacobs_2016, Wang_2018}. Increasing turbulence is is one mechanism by which collectors could affect settling rates.

@@

We used a model adapted from \cite{Fauria_2015} to estimate collection efficiency based on exponential decay of suspended sediment concentration, after introducing sediment in a single, initial dose. The key equation in the exponential decay model is \[\frac{d\bar{\phi}}{dt} = -[\frac{Cv_s}{h}(1-E_r) + \eta^{\prime}ud_cI_c]\bar{\phi} = -k\bar{\phi}\,.\] The two summands in brackets represent the net decrease of average suspended settlement concentration ($\bar{\phi}$) due to settling and collection respectively. $C$ is a constant based on the shape of the vertical sediment profile, which is assumed to follow the Rouse equation; $v_s$ is the settling velocity; $h$ is water depth, which equals the collector height for emergent collectors; and $E_r$ is the rate of entrainment (i.e., uptake of particle from the channel bed).
